---
title: "feather river redd survey locations naming qc - adding names to data"
date: "2024-12-09"
rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
library(readxl)
library(sf)
library(openxlsx)
library(readxl)
```
The goal of this markdown is to clean up the location names from Feather River redd survey data.

Summary of work that has been done so far:

  * redd_qc_2014-2023 - mainly exploratory: 
    *  raw data is pulled in, negative velocity values are set to NA
    * pot_depth_m >= 40 is set to NA
    * "type" is removed
    
  * mapping_qc: looks into latitude longitude and identifies issues. 
    * Data file input: "data/archive/all_redd_data.csv" 
  * clean_locations: fixed latitude/longitude provided by Chris Cook, these changes address the issues found on mapping_qc. 
    * Data file input: "data/archive/all_redd_data.csv". 
    * Data file output(s): "data-raw/redd_observations_clean_locations.csv" and "data/redd_observations.csv" (`location`, `survey_wk` and `file_number` removed)
  
```{r, include=FALSE}
redd_observations <- read.csv("data-raw/redd_observations_clean_locations.csv")
```

Looking into unique location names
```{r}

```

