---
title: "feather river redd survey locations naming qc - adding names to data"
date: "2024-12-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
library(readxl)
library(sf)
library(leaflet)
library(openxlsx)
library(readxl)
library(stringr)
```

The goal of this markdown is to clean up the location names from Feather River redd survey data.

Summary of work that has been done so far:

  * *redd_qc_2014-2023* - mainly exploratory: 
    * raw data is pulled in, negative velocity values are set to NA
    * pot_depth_m >= 40 is set to NA
    * "type" is removed
    * 2023 number_redds set all to 1 
    * note that 7 data entries in the raw data that had "0" in number_redds, those were set to 1
    
  * *mapping_qc*: looks into latitude longitude and identifies issues. 
    * Data file input: "data/archive/all_redd_data.csv" 
  * *clean_locations*: fixed latitude/longitude provided by Chris Cook, these changes address the issues found on mapping_qc. 
    * Data file input: "data/archive/all_redd_data.csv". 
    * Data file output(s): "data-raw/redd_observations_clean_locations.csv" and "data/redd_observations.csv" (`location`, `survey_wk` and `file_number` removed)
  
```{r, include=FALSE}
redd_observations <- read.csv("data-raw/redd_observations_clean_locations.csv")
```

### Looking into unique location names

  * There are 81 unique names
  
```{r}
unique(redd_observations$location)
```

```{r, include=FALSE}
# removing spaces
redd_observations$location <- gsub(" ", "", redd_observations$location)

# Display the modified vector
print(unique(redd_observations$location))

```


Survey location names - 
The following names below are survey locations names provided by Feather River team. These location names are what they have on their reports and how they categorize the locations where surveys are conducted. Redd data locations will be matched to these names

```{r, echo=FALSE, message=FALSE}
# Names from survey location provided by Chris Cook
# Great Western, Palm, Top of Hatchery, G95 Main and Lower Vance East are not included on locations provided but we asked Chris and we should keep them
reference_locations <- c(
  "UpperHatchery", "UpperCottonwood", "Hatchery", "Cottonwood",
  "TopofAuditorium", "MoesSideChannel", "UpperAuditorium", "MiddleAuditorium",
  "LowerAuditorium", "HatcheryPipe", "LowerTableMountain", "TableMountain",
  "BelowLowerAuditorium", "Bedrock", "TrailerPark", "UpperMathews",
  "Mathews", "UpperRobinson", "LowerRobinson", "Steep",
  "SteepSideChannel", "Weir", "Eye", "EyeSideChannel",
  "Gateway", "Aleck", "HatcherySideChannel", "VanceEast",
  "VanceWest", "BigHoleEast", "BigHoleWest", "BelowBigHoleEast",
  "G95WestSideChannel", "G95EastSideChannelTop", "G95EastSideChannelBottom", "UpperHour",
  "LowerHour", "Keister", "Goose", "Big",
  "LowerBig", "BigBar", "UpperMcFarland", "LowerMcFarland",
  "Developing", "GreatWestern", "Palm", "TopofHatchery", "G95main", "LowerVanceEast"
)

reference_locations <- tolower(reference_locations)
table_data <- data.frame(
  ID = seq_along(reference_locations), 
  Name = reference_locations
)
table_data <- table_data[order(table_data$Name), ]

# Generate the sorted table
knitr::kable(table_data, col.names = c("ID", "Location Name"))
```

Identifying redd locations that do not match survey location names

```{r, echo=FALSE}
invalid_locations <- redd_observations$location[!redd_observations$location %in% reference_locations]

# Display invalid locations
unique_invalid_locations <- unique(invalid_locations)
print(unique_invalid_locations)
```

```{r, include=FALSE}
# renaming those that are just miss spelled 
redd_observations_clean_1 <- redd_observations |> 
  mutate(location = case_when(
    location == "aleckriffle" ~ "aleck",
    location == "bedrockriffle" ~ "bedrock",
    location == "bigriffle" | location ==	"bigbarriffle" ~	"bigbar",
    location == "g-95eastsidechannelbottom"	| location == "g95eastbottom"	~ "g95eastsidechannelbottom",
    location == "g-95easttop" | location == "g95easttop"	~ "g95eastsidechanneltop",
    location == "g-95westsidechannelbottom" | location ==	"g95west" | location ==	"g95westbottom" ~	"g95westsidechannel",
    location == "gooseriffle"	~	"goose",
    location == "hatcheryriffle" ~ "hatchery", # unsure relationship
    location == "hatcherysidechannelupper" ~	"hatcherysidechannel",
    location == "lowerbigriffle" ~	"lowerbig",
    location == "matthews" ~	"mathews",
    location == "midauditorium"| location == "midhour" ~	"middleauditorium", # unsure about midhour relationship
    location == "moe'ssidechannel" ~ "moessidechannel",
    location == "steepriffle"	~	"steep", # unsure
    location == "steepside" ~ "steepsidechannel", #unsure
    location == "topofaudtorium" ~ "topofauditorium",
    location == "upperhatcheryriffle" ~	"upperhatchery",
    location == "uppermatthews" ~	"uppermathews",
    location == "weirriffle"	~	"weir",
    location == "developingriffle" ~ "developing",
    location == "gatewaysidechannel" | location == "gatewaymainchannel" ~ "gateway",
    location == "upperbigholeeast" ~ "bigholeeast",
    location == "g-95main" ~ "g95main",
    location == "palmriffle" ~ "palm",
    TRUE ~ location)) 

print(unique(redd_observations_clean_1$location))
```

Redd location names that do not match survey locations, even after naming cleaning

```{r, echo=FALSE}
# checking for the locations that still dont match reference names, after cleaning above
print(unique(redd_observations_clean_1$location[!redd_observations_clean_1$location %in% reference_locations]))
```


```{r, include=FALSE}
names_for_check <- redd_observations_clean_1 |> 
   filter(!location %in% reference_locations)  # Keep only the rows with inconsistent location names
# not_consistent_names <- st_as_sf(not_consistent_names, coords = c("longitude", "latitude"), crs = 4326)
```

Reference map to check where redds are located geographically

```{r echo=FALSE, warning=FALSE}
# pulling layers from survey location map for reference
gdb_path <- "data-raw/shapefiles/b105_redd_survey_locations_maps_2021.gdb"
layers <- st_layers(gdb_path)$name

name_check_map <- leaflet() |> 
  addTiles() |> 
  setView(lng = -121.5695, lat = 39.3665, zoom = 10)

for (layer_name in layers) {
  shapefile <- st_read(dsn = gdb_path, layer = layer_name)
  shapefile <- st_transform(shapefile, crs = 4326) |> 
    st_zm(drop = TRUE, what = "ZM")
  
name_check_map <- name_check_map |> 
    addPolygons(
      data = shapefile,
      fillColor = "lightblue",  
      color = "darkblue",
      weight = 2,               
      opacity = 1.0,            
      fillOpacity = 0.5,        
      group = layer_name,
      popup = paste("Layer Name: ", layer_name)
    )
}

locations <- unique(names_for_check$location)
name_check_map <- name_check_map |> 
  addCircleMarkers(data = names_for_check, 
                   ~longitude, ~latitude, 
                   label = ~paste(location),
                   popup = ~paste("Location:", location, "<br>Date:", date, "<br>Latitude:", latitude, "<br>Longitude:",
                                  longitude, "<br>Survey Week:", survey_wk),
                   group = ~location,
                   radius = 2,  
                   color = "red")

name_check_map <- name_check_map |> 
  addLayersControl(
    overlayGroups = c(locations), # Add chek boxes for the location names 
    options = layersControlOptions(collapsed = FALSE)
  )

# Display the map
name_check_map
```

### Notes / observations form map:

  * Great Western, Palm, Top of Hatchery, G95 Main and Lower Vance East are not included on location shapefiles provided, but we asked Chris and we should keep them
  * hour: most points are located near "LowerHour", however two are within "UpperHour"
  * palm / palmriffle are not within any survey location - it was not modified
  * g95main: There are two points that look like outliers from other points - we asked Chris and poins have been updated accordingly below
  * lowervanceeast: 3 points are within "BigHoleEast" (these were modified to BigHoleEast), all others are off any survey location - south of "VanceEast", north of "BigHoleEast"
  * gatewaymainchannel: located on survey location Gateway, modified
  * upperbigholeeast: modified to BigHoleEast
  * moe'sditch: modified 
  * upper mathews: 
  
Modifications - all points that fall within a survey location will be modified to match to that name

```{r, include=FALSE}
redd_observations_clean_2 <- redd_observations_clean_1 |>
  mutate(location = case_when(
     location == "hour" & date == "2015-11-19" & latitude == "39.4225537735696" & longitude == "-121.624800693264" ~ "upperhour",
     location == "hour" & date == "2015-11-19" & latitude == "39.4224821992187" & longitude == "-121.624878645819" ~ "upperhour",
     location == "hour" ~ "lowerhour",
     location == "upperhoureast" & date == "2014-10-09" & latitude == "39.4166671193809" & longitude == "-121.627212180498"
    ~ "lowerhour",
    location == "upperhoureast" & date == "2014-10-09" & latitude == "39.4159674094301" & longitude == "-121.627254551989" 
    ~ "lowerhour",
    location == "developingriffle" ~ "developing",
    location == "moe'sditch" & !(latitude == "39.5156950186919" & longitude == "-121.559350861003") ~ "moessidechannel",
    location == "moe'sditch" & latitude == "39.5156950186919" & longitude == "-121.559350861003" ~ "middleauditorium",
    location == "lowervanceeast" & date == "2017-11-14" & latitude == "39.439059494603" & longitude == "-121.635427259305" ~ "bigholeeast",
    location == "lowervanceeast" & date == "2017-11-14" & latitude == "39.4388562088935" & longitude == "-121.635448923638" ~ "bigholeeast",
    location == "lowervanceeast" & date == "2017-11-14" & latitude == "39.4388504794907" & longitude == "-121.635372994652" ~ "bigholeeast",
    location == "g95main" & date == "2015-11-19" & latitude == "39.4394792142928" & longitude == "-121.635254377528" ~ "bigholeeast",
    location == "g95main" & date == "2015-11-19" & latitude == "39.4371081082104" & longitude == "-121.636483707754" ~  "bigholeeast",
    location == "uppermathews" & date == "2023-10-25" & latitude == "39.5157891" & longitude == "-121.55643006" ~ "upperauditorium", # TODO from here below were modified on 01/08/25 - might want to confirm with Chris that modifications are ok. Names were changed based on their actual location
    location == "uppermathews" & date == "2023-10-25" & latitude == "39.51585525" & longitude == "-121.55658786" ~ "upperauditorium",
    location == "uppermathews" & date == "2023-10-25" & latitude == "39.51600506" & longitude == "-121.55655445" ~ "upperauditorium",
    location == "uppermathews" & date == "2023-10-25" & latitude == "39.51589657" & longitude == "-121.55733036" ~ "upperauditorium",
    location == "uppermathews" & date == "2023-10-19" & latitude == "39.51572598" & longitude == "-121.5576469" ~ "upperauditorium",
    location == "uppermathews" & date == "2023-10-19" & latitude == "39.51570684" & longitude == "-121.55787447" ~ "upperauditorium",
    location == "trailerpark" & date == "2023-09-26" & latitude == "39.51600036" & longitude == "-121.55916329" ~ "moessidechannel",
    location == "tablemountain" & date == "2022-11-22" & latitude == "39.496410384211" & longitude == "-121.579623191602" ~ "trailerpark",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4176998620939" & longitude == "-121.627223819039" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.417653889353" & longitude == "-121.627343452105" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4174661545398" & longitude == "-121.62741308081" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4174294735085" & longitude == "-121.627411511883" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4173586922004" & longitude == "-121.627554538371" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.41572855345" & longitude == "-121.626946544251" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4157329027447" & longitude == "-121.627060052528" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4156719809222" & longitude == "-121.627129956771" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4156289839513" & longitude == "-121.627133750753" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4156255231564" & longitude == "-121.627261165499" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4098841591031" & longitude == "-121.625766739867" ~ "lowerhour",
    location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4192277254991" & longitude == "-121.626608190798" ~ "lowerhour",
    TRUE ~ location
  ))


# pending points to ask Chris - 
# pending <- redd_observations_clean_2 |> 
#   filter(location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4192277254991" & longitude == "-121.626608190798" |
#          location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4098841591031" & longitude == "-121.625766739867") |> 
#   glimpse()


print(unique(redd_observations_clean_2$location))

print(unique(redd_observations_clean_2$location[!redd_observations_clean_2$location %in% reference_locations]))
# all redd locations now fall within reference locations!
```

```{r, include=FALSE}
#pending as of 01/08/25 --- RESOLVED
# redd_questions <- redd_observations_clean_1 |> 
#   filter(location == "uppermathews" & date == "2023-10-25" & latitude == "39.5157891" & longitude == "-121.55643006" |
#     location == "uppermathews" & date == "2023-10-25" & latitude == "39.51585525" & longitude == "-121.55658786" |
#     location == "uppermathews" & date == "2023-10-25" & latitude == "39.51600506" & longitude == "-121.55655445" |
#     location == "uppermathews" & date == "2023-10-25" & latitude == "39.51589657" & longitude == "-121.55733036" |
#     location == "uppermathews" & date == "2023-10-19" & latitude == "39.51572598" & longitude == "-121.5576469" | 
#     location == "uppermathews" & date == "2023-10-19" & latitude == "39.51570684" & longitude == "-121.55787447"| 
#     location == "trailerpark" & date == "2023-09-26" & latitude == "39.51600036" & longitude == "-121.55916329" | 
#     location == "tablemountain" & date == "2022-11-22" & latitude == "39.496410384211" & longitude == "-121.579623191602" | 
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4176998620939" & longitude == "-121.627223819039" | 
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.417653889353" & longitude == "-121.627343452105" |
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4174661545398" & longitude == "-39.4174661545398" | 
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4174294735085" & longitude == "-121.627411511883" | 
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4173586922004" & longitude == "-121.627554538371" | 
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.41572855345" & longitude == "-121.626946544251" | 
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4157329027447" & longitude == "-121.627060052528" | 
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4156719809222" & longitude == "-121.627129956771"| 
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4156289839513" & longitude == "-121.627133750753" | 
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4156255231564" & longitude == "-121.627261165499" |
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4174661545398" & longitude == "-121.627261165499" |
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4192277254991" & longitude == "-121.626608190798" |
#     location == "middleauditorium" & date == "2014-11-07" & latitude == "39.4098841591031" & longitude == "-121.625766739867") |>
#   glimpse()
# 
# write.csv(redd_questions, "data-raw/redd_questions_0125.csv")
```

Listing redd location names that remain unmatched to any survey locations (should be 0)

```{r, echo=FALSE}
# another check
names_for_check_2 <- redd_observations_clean_2 |> 
   filter(!location %in% reference_locations) 
  

unique(names_for_check_2$location)
```


```{r, include=FALSE}
redd_observations_wnames <- redd_observations_clean_2 |> 
  mutate(date = case_when(
    str_detect(date, "/") ~ as.Date(date, format = "%m/%d/%Y"), # fix dates
    TRUE ~ as.Date(date))) |> 
  select(-file_number) |>  # keeping survey week and location
  view() 
  
# save clean data
write.csv(redd_observations_wnames, "data/redd_observations_wnames.csv", row.names = FALSE)
```


