---
title: "feather river redd data exploration"
output: html_document
date: "2024-06-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
```

Goal of this document is to revisit Feather River redd data and understand the following:

- what data are collected?
- what is the quality of the data?
- what are the data limitations?


```{r, include = F}
# get standard data from cloud
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_redd.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data-raw", "feather-redd.csv"),
               overwrite = TRUE)

feather_redd_raw <- read_csv("data-raw/feather-redd.csv")
```

# What data are collected

```{r}
feather_redd_raw |> glimpse()
```

## Questions

- Need to better understand `type` field. Do "area" measurements have a shapefile associated with them? Is "point" meant to communicate that a specific point measurement was taken for a redd?
- We need to QC the latitude/longitude - looks like there is a decimal place missing.
- Why is the `redd_count` not always 1? If greater than 1 does that mean that redd specific measurements like depth are not collected?
- Does salmon count mean that salmon were observed on redd? No species or run differentiation?
- Where are depth measurements taken (just depth, not pot depth)?
- Does `pot_depth` refer to the depth of the pit?
- How are velocity measurements taken? Collected right above redd?
- Where is width measured?

# Data quality

* Pull from JPE Rmd to summarize NAs, synthesize questions/concerns about data
* Create a map of latitude/longitude to QC

# Data limitations

Summary of how we may be able to use the data and limitations for different uses