---
title: "feather river redd data exploration"
date: "2024-08-27"
output: 
  html_document:
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(dplyr)
library(leaflet)
library(RColorBrewer)
library(viridis)
library(readxl)
library(janitor)
library(sf)
library(openxlsx)
library(readxl)
```

Goal of this document is to review the Feather River redd data and understand the following:

- what data are collected?
- what is the quality of the data?
- what are the data limitations?
- how are these data useful to Science Plan hypotheses?

```{r, include=FALSE, warning=FALSE}
redd_2014 <- readxl::read_excel(here::here("data-raw", "2014-2023 Chinook Redd Survey Data", "2014_Chinook_Redd_Survey_Data.xlsx")) |> 
  clean_names() |> 
  mutate(date = convertToDate(date)) |> 
  slice(-1912) |>  #removing the totals column
  glimpse()


# Read the data

  # mutate(Date = as.Date.POSIXct(as.numeric(Date))) |> 
  # glimpse()

redd_2015 <- readxl::read_xlsx(here::here("data-raw","2014-2023 Chinook Redd Survey Data","2015_Chinook_Redd_Survey_Data.xlsx"), range = cell_rows(1:2353)) |> 
  clean_names() |> 
  glimpse()

redd_2016 <- readxl::read_xlsx(here::here("data-raw","2014-2023 Chinook Redd Survey Data","2016_Chinook_Redd_Survey_Data.xlsx"), range = cell_rows(1:1571)) |> 
  clean_names() |> 
  glimpse()

redd_2017 <- readxl::read_xlsx(here::here("data-raw","2014-2023 Chinook Redd Survey Data","2017_Chinook_Redd_Survey_Data.xlsx"), range = cell_rows(1:2718)) |> 
  clean_names() |> 
  glimpse()

redd_2018 <- readxl::read_xlsx(here::here("data-raw","2014-2023 Chinook Redd Survey Data","2018_Chinook_Redd_Survey_Data.xlsx"), range = cell_rows(1:4170)) |> 
  clean_names() |> 
  glimpse()

redd_2019 <- readxl::read_xlsx(here::here("data-raw","2014-2023 Chinook Redd Survey Data","2019_Chinook_Redd_Survey_Data.xlsx"), range = cell_rows(1:5049)) |> 
  clean_names() |> 
  glimpse()

redd_2020 <- readxl::read_xlsx(here::here("data-raw","2014-2023 Chinook Redd Survey Data","2020_Chinook_Redd_Survey_Data.xlsx"), range = cell_rows(1:5432)) |> 
  clean_names() |> 
  glimpse()

redd_2021 <- readxl::read_xlsx(here::here("data-raw","2014-2023 Chinook Redd Survey Data","2021_Chinook_Redd_Survey_Data.xlsx"), range = cell_rows(1:2595)) |> 
  clean_names() |> 
  glimpse()

redd_2022 <- readxl::read_xlsx(here::here("data-raw","2014-2023 Chinook Redd Survey Data","2022_Chinook_Redd_Survey_Data.xlsx"), range = cell_rows(1:3762)) |> 
  clean_names() |> 
  glimpse()
```


```{r, include=FALSE, warning=FALSE}
# this year's fields look different
redd_2023 <- readxl::read_xlsx(here::here("data-raw","2014-2023 Chinook Redd Survey Data","2023_Chinook_Redd_Survey_Data.xlsx")) |>
  clean_names() |>
  glimpse()

#testing with just the first tab of excel book
redd_2023_clean <- redd_2023 |> 
  mutate(date =  mdy_hm(date_time),
         survey_wk = str_extract(x2, "\\d+-\\d+"),
         location = riffle_name,
         file_number = NA,
         type = NA,
         number_redds = number_of_fish_on_redd,
         number_salmon = NA,
         latitude_m_n = latitude_wgs_84,
         longitude_m_e = longitude_wgs_84,
         depth_m  = water_depth_m,
         pot_depth_m = NA,
         velocity_m_s = velocity_m_sec,
         percent_fines = percent_sand_0_0625_2mm, #should we also include % Silt <0.062mm
         percent_small = percent_gravel_2_64mm,
         percent_med = percent_pebble_64_128,
         percent_large = percent_cobble_128_256mm,
         percent_boulder = percent_boulder_256mm,
         longitude_n_e = NA) |> 
  select(date, survey_wk, location, file_number, type, number_redds, number_salmon, latitude_m_n, longitude_m_e, depth_m, pot_depth_m, velocity_m_s, percent_fines, percent_small, percent_med, percent_large, percent_boulder, redd_width_m, redd_length_m, longitude_n_e) |> 
  glimpse()

```
field relationship 2023 to data from other years 

date - date/time 
survey_wk - col B (no name)
location - Riffle name
file_number - NA
type - NA
number_redds - Number of fish on Redd
number_salmon - NA
latitude_m_n - Latitude (WGS 84)
longitude_m_e - Longitude (WGS 84)
depth_m  - water depth (m)
pot_depth_m - ?
velocity_m_s - Velocity (m/sec.)
percent_fines - % Sand 0.0625-2mm #should we also include % Silt <0.062mm
percent_small - % Gravel 2-64mm
percent_med - % Pebble 64-128
percent_large - % Cobble 128-256mm
percent_boulder - % Boulder >256mm
redd_width_m - Redd Width (m)
redd_length_m - Redd Length (m)

Unmatched fields:
redd #
Elevation (ft)
Accuracy (ft)
Subsample
Boat Point 
Comment 
User


Combining all years into one object

```{r}
combined_redd <- bind_rows(redd_2014, redd_2015, redd_2016, redd_2017, redd_2018, redd_2019, redd_2020, redd_2021, redd_2022) |> 
  glimpse()
```

```{r}
str(combined_redd)
```

### Data exploration {.tabset}

Exploring fields that have data consistently

#### date
 
```{r}
#date range
range(combined_redd$date, na.rm = TRUE)
```


#### location

  * Location clean up is needed. There are some names that appear less than 5 times. We could check if they correspond to another location name, and it is spelled differently

```{r}
#locations
table(combined_redd$location) #there are 121 unique location names
```

**Locations surveyed less than 5 times**

For locations that are being surveyed less than five times, we might want to verify if they belong to other location area names. Otherwise, these sites might not have enough data to reliably analyze over time.

```{r}
#plotting those locations that appear less than 5 times
combined_redd |> 
  group_by(location) |> 
  filter(n() < 5) |> 
  ungroup() |> 
  count(location) |> 
  ggplot(aes(x = reorder(location, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  labs(title = "Count of Locations with Less Than 5 Observations",
       x = "Location",
       y = "Number of Data Records") +
  theme_minimal()

```

**Frequent locations and years surveyed**

```{r, warning=FALSE}
location_frequent <- combined_redd |> 
  mutate(year = year(date)) |> 
  select(location, year) |> 
  group_by(location, year) |> 
  distinct() |> 
  ungroup() |> 
  group_by(location) |> 
  tally() |> 
  filter(n > 5)

combined_redd |> 
  mutate(year = as.factor(year(date))) |> 
  select(location, year) |> 
  group_by(location, year) |> 
  distinct() |> 
  filter(location %in% location_frequent$location) |> 
  ggplot(aes(x = year, y = location, color = year)) +
  geom_point(size = 2, alpha = 0.7) +
  labs(title = "Distinct Locations by Year",
       x = "Year",
       y = "Year",
       color = "Location") +  
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text = element_text(size = 8),
        legend.position = "none")
  
```


#### type

- note that there is one type record that is 3761, with no date associated, is this an error?

**Observations by type and date**

```{r, warning=FALSE}
#type
unique(combined_redd$type)

ggplot(combined_redd, aes(x = date, y = type, color = type)) +
  geom_point(size = 2, alpha = 0.7) +
  labs(title = "Types Over Time",
       x = "Date",
       y = "Type",
       color = "Type") +
   scale_color_viridis_d(na.value = "grey50") +  
  theme_minimal() +
# Use Viridis color scale
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text = element_text(size = 10))
```


**Count of observations by type**

```{r}
combined_redd |> 
  group_by(type) |> 
  summarise(n = n()) |> 
  ggplot(aes(x = type, y = n, fill = type)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = n), vjust = -0.5, size = 3.5) +
  scale_fill_viridis_d() +  
  labs(title = "Count of Observations by Type",
       x = "Type",
       y = "Count") +
  theme_minimal() 
```

#### redd_count

**Redd counts by year**

```{r}
range(combined_redd$number_of_redds, na.rm = TRUE)

combined_redd <- combined_redd %>%
  mutate(date = as.Date(date)) 

#TODO clean up this plot 

# ggplot(combined_redd, aes(x = date, y = number_of_redds)) + 
#   geom_col() +
#   facet_wrap(~year(date), scales = "free") +
#   scale_x_date(labels = date_format("%b"), date_breaks = "1 month") +
#   labs(title = "Redd Counts Over Time") +  
#   theme_minimal() +
#   theme(
#     strip.background = element_rect(fill = "lightgrey", color = "grey50"),
#     plot.title = element_text(hjust = 0.5) 
#   )

```


#### salmon_count

**Salmon counts by year**

```{r, warning=FALSE}
range(combined_redd$number_salmon, na.rm = TRUE)

combined_redd |> 
  filter(is.na(date)==FALSE) |>  
  ggplot(aes(x = date, y = number_salmon)) + 
  geom_col() +
  facet_wrap(~year(date), scales = "free") +
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") +
  labs(title = "Salmon Counts Over Time") +  
  theme_minimal() +
  theme(
    strip.background = element_rect(fill = "lightgrey", color = "grey50"),
    plot.title = element_text(hjust = 0.5) 
  )
```

Latitude and longitude needs to be transformed to UTM
#TODO

```{r, include = F}
# # Filter out rows with NA in either latitude or longitude
# combined_redd_clean <- combined_redd |> 
#   filter(!is.na(longitude_m_e) & !is.na(latitude_m_n))
# 
# # Create an sf object from your data
# combined_redd_sf <- combined_redd_clean |> 
#   mutate(
#     longitude_m_e = ifelse(is.na(longitude_m_e), NA, longitude_m_e),
#     latitude_m_n = ifelse(is.na(latitude_m_n), NA, latitude_m_n)) |> 
#   st_as_sf(coords = c("longitude_m_e", "latitude_m_n"), crs = 4326)
# 
# # Transform to UTM (replace "32633" with the appropriate UTM EPSG code for your zone)
# combined_redd_utm <- st_transform(combined_redd_sf, crs = 32611)

```
